Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
Rewording
minors

In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
We maintain all nodes of a component always collocated in the same cluster~\cite{repartition-disc}.
When we  move a~node, we move the whole component that contains it.
We maintain a balanced partition of our components,
a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.

@@

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	\RM collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter.
	I.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to a pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	Note that \OPT cannot have both pairs collocated at the same time.
	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume  \OPT has one pair,
	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$,
	separating $\{u,v\}$.
	We define times $t'' < t' < m$ similarly as in the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\mathit{OPT} (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\OPTM ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
	and  
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by \RM,
	and for final costs at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in T } \mathit{OPT} (R_t) +
	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

@@

\maciek{"similarly as in the former" "\textbf{similarly to the former}"}

@@

\maciek{OPT's cost formula: make it one line by removing the expression with minimums. Move minimums to the paragraph text, by immediatelly expanding minimums (two steps in paragraph, instead of in the formula)}

>>>>

In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
We keep all nodes of each component collocated in the same cluster, similar to Avin et al.~\cite{repartition-disc}.
When we  move a~node, we move the whole component that contains it.
We maintain a balanced partition of our components,
a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.

@@

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t$,
	\RM collocates the pair $\set{u,v}$.
	This means,
	the value of $C_{\set{u,v}}$ at $t$,
	denoted by $C^t_{\set{u,v}}$, 
	reaches $\lambda$ immediately before \RM resets the counter,
	i.e.,
	$ C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to a pair $\set{x,y}$ that arrive
	during $[t_1,t_2]$ by $\sigma_{\set{x,y}}[t_1,t_2]$,
	and its overall cost to \OPT by
	$\mathit{OPT} (\sigma_{\{u,v\}}[t_1,t_2])$.
	We may use $\sigma_{\set{x,y}}$ whenever
	 the interval $[t_1,t_2]$ is clear from the context.
	%
	%Else, this is not the first time and
	%Alg collocates $\{u,v\}$ also some time prior to $t$.

@@

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving 	and the moving cost is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$,
	 then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	If \OPT has both pairs separated during their respective intervals,
	then obviously it pays $2\lambda$ during those intervals.
	Note that \OPT cannot have both pairs collocated at the same time.
%	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume \OPT has one of the pairs,
	e.g.~$\{u,v\}$,
	 collocated already prior its respective interval, $(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$,
	separating $\{u,v\}$.
	We define times $t'' < t' < m$ similarly to the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\mathit{OPT} (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\OPTM ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
	and  
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by \RM,
	and for final requests at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in T } \mathit{OPT} (R_t) +
	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

@@

@@

\maciek{OPT's cost formula: make it one line by removing the expression with minimums. Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula)}
